\chapter{CONCLUSION}

The objectives of recognizing a hand-drawn circuit diagram, transforming it into an editable schematic having simulation capabilities, and analyzing said circuit with LLM-powered Q\&A have been successfully achieved. Functional pipelines for each aspect of the project have been designed and implemented effectively. For component detection, YOLOv7 has achieved a mAP@0.5 of 95\% and for text recognition, TrOCR has achieved accuracy of 84.5\%. The wire detection algorithm leverages junction-based geometric inference with adaptive collinear merging, enabling robust extraction of circuit connectivity. The system processes hand-drawn circuit images and extracts components with their values and connections, effectively digitizing a circuit. The project has successfully addressed a significant gap in educational circuit analysis tools by making circuit simulation easier than ever.

Apart from technical performance, the project also demonstrates successful full-stack integration with a React.js frontend, a FastAPI backend, and an integration with CircuitJS for simulation of the circuit. This modular architecture ensures ease of maintenance and potential extensibility, establishing a solid foundation for future enhancements.

This project also serves as a vehicle for research and experimentation for its authors. The study into different versions of YOLO for component detection, different OCR models for text recognition, as well as multiple ways to solve wire connectivity issues has not only brought this project as close to the best as it could have been, but also infused the authors with the knowledge and technical know-how regarding a variety of topics and cutting-edge scientific research. Circuitron represents a meaningful step towards democratizing ciruit education through the elimination of friction between ideation and analysis. It establishes a replicable and of course, improvable, design pattern for the broader electronics education community. Its potential to improve student engagement while reducing cognitive load makes it a valuable contribution to electrical and electronics education, where traditional circuit simulation tools are expensive or complex. 
 
